---
title: "Analysis - Appendix B"
author: "Alexandre Kobrin - Bachelor Thesis"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
subtitle: Documentation of Analysis in R
---

## Loading packages and data

```{r, warnings= FALSE }

# Load the necessary packages
library(tidyverse)
library(car)
library(knitr)
library(glmnet)
library(performance)
library(ggplot2)
library(dplyr)
library(glmnet)
library(caret)
library(readr)
library(latex2exp)
```

Easiest way to have the data in this notebook is to import the database, however bellow is an option with a manual path that can be commented out.

```{r}
# Set working directory 
#knitr::opts_knit$set(root.dir = "path to file")

# Importing data
rent_data <- read.csv('data/rents_S3_Done.csv')

#rent_data <- rents_S3_Done
```



This is a quick small overview of the data to check for any anomalies. To see more the table for each variable can be commented and un-commented.
```{r}

table(rent_data$a_nb_rooms)
table(rent_data$a_netm_mon)
table(rent_data$age)
table(rent_data$a_sicht)
table(rent_data$a_balkon)
table(rent_data$region)
table(rent_data$a_zip_2)
table(rent_data$a_kat_o_2)

```

## Preparing data for regression

Certain variables are transformed to categorical variables. The mutate() function is used to create new variables based on the existing variables, and the case_when() function define categories for the numerical values. It assignes descriptive labels to each category within the variables a_sicht (view), a_balkon (balcony), and region. These variables are considered nominal categorical variables, as their categories only serve to distinguish between different groups or attributes without any hierarchical structure.
Next, the mutate_if() function is used to convert all character variables to factor variables, ensuring that the categorical variables are treated as factors in the analysis. This allows allows R to create the necessary dummy variables for each category.
Lastly, the a_nb_rooms variable is treated as an ordered categorical variable. The rooms variable is ordered as unlike the other categorical variables, rooms can be ranked from fewer to more rooms, which implies a progression in the data.

```{r}

#assigning the different numerical values to their categories 
rent_data <- rent_data %>%
  mutate(
    a_sicht = case_when(
      a_sicht == 0 ~ 'no_view',
      a_sicht == 1 ~ 'in_general',
      a_sicht == 2 ~ 'to_the_mountains',
      a_sicht == 3 ~ 'on_the_lake',
      a_sicht == 4 ~ 'Mountains_and_lake',
      a_sicht == 5 ~ 'Not_lake_and_mountain'),
    a_balkon = case_when(a_balkon == 1 ~ 'no balcony',
                         a_balkon == 2 ~ 'general',
                         a_balkon == 3 ~ 'small balcony',
                         a_balkon == 4 ~ 'big balcony',
                         a_balkon == 5 ~ 'terasse',
                         a_balkon == 6 ~ 'roof terasse'),
    region = case_when(region == 0 ~ 'german',
                       region == 1 ~ 'french',
                       region == 2 ~ 'italian'),
    a_kat_o_2 = case_when(a_kat_o_2 == "AP" ~ 'apartment',
                       a_kat_o_2 == "HO" ~ 'house')
  )

#convert all character variables to factors
rent_data <- rent_data %>%
  mutate_if(is.character, as.factor)

#set number of rooms as an ordered factor variable
sort(unique(rent_data$a_nb_rooms))
rent_data$a_nb_rooms<-factor(rent_data$a_nb_rooms,levels =c(sort(unique(rent_data$a_nb_rooms))),ordered = T )
rent_data$a_nb_rooms<-as.character(rent_data$a_nb_rooms)
unique(rent_data$a_nb_rooms)
glimpse(rent_data)

```

## Model 1 - No interactions

Building multiple linear regression model with all the variables included but no interaction term.

```{r}
options(scipen = 999)
no_interaction_model <- lm(a_netm_mon ~ a_surface_living + age + a_sicht + a_balkon + region + a_zip_2 + a_nb_rooms + a_kat_o_2, data=rent_data)
summary(no_interaction_model)

```

Variance Inflation Factor test on model 1, if there are variables that has a greater value then 5 they could be subject to multicolinearity issues.

```{r}
vif(no_interaction_model)
```

## Checking for Interactions

The following plots are visualizations of the interaction terms between tdifferent variables and the region variable, showing how the relationship between them and net monthly rent (a_netm_mon) varies across different regions. 

The geom_smooth() function is used to fit linear regression models for each region, with the method = 'lm' argument specifying the use of linear models. The se = FALSE argument removes the shaded confidence intervals around the regression lines, resulting in a cleaner and more focused visualization of the interaction effect.


#### Interaction between Living Surface and Region

```{r,warnings= FALSE, message=FALSE}
ggplot(data = rent_data, aes(x = a_surface_living, y = a_netm_mon, color = region)) +
  geom_smooth(method = 'lm', se = FALSE, aes(linetype = region), size = 1.2) +
  scale_color_manual(values = c("black", "black", "black")) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  theme_classic() +
  guides(linetype = FALSE)


```


#### Interaction between Age of building and region and Region

```{r,warnings= FALSE, message=FALSE}
ggplot(data = rent_data, aes(x = age, y = a_netm_mon, color = region)) +
  geom_smooth(method = 'lm', se = FALSE, aes(linetype = region), size = 1.2) +
  scale_color_manual(values = c("black", "black", "black")) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  theme_classic() +
  guides(linetype = FALSE)


```

#### Interaction between number of rooms and Region

```{r,fig.width=15, fig.height=10, warnings= FALSE, message=FALSE}
interaction.plot(x.factor = rent_data$region,
                 trace.factor = rent_data$a_nb_rooms,
                 response = rent_data$a_netm_mon, fun = mean,
                 col = ifelse(rent_data$a_nb_rooms == 1, "black", "black"),
                 lwd = 1.2, lty = 1:3, type = "b",
                 xlab = "Region", ylab = "Monthly Net Rent")

```

#### Interaction between view and Region

```{r, fig.width=15, fig.height=10}
interaction.plot(x.factor = rent_data$region,
                 trace.factor = rent_data$a_sicht,
                 response = rent_data$a_netm_mon, fun = mean)

```

#### Interaction between Balcony and Region

```{r,fig.width=15, fig.height=10}
interaction.plot(x.factor = rent_data$region,
                 trace.factor = rent_data$a_balkon,
                 response = rent_data$a_netm_mon, fun = mean)

#Extraction of exact values since the interaction is small 
x<-rent_data %>% 
  group_by(region,a_balkon) %>% 
  summarise(a_netm_mon=mean(a_netm_mon,na.rm=T))
kable(x, format = "markdown")
```

## Model 1.1.2 - Interactions

Building multiple linear regression model with all the variables and relevant interaction term.

```{r}
options(scipen = 999)
interaction_model <- lm(a_netm_mon ~ a_surface_living + age + a_sicht + a_balkon + region + a_zip_2 + a_nb_rooms + a_kat_o_2 + region:a_surface_living + region:a_nb_rooms + region:a_sicht + region:a_balkon, data=rent_data)
summary(interaction_model)

```

## Verifying regression assumptions

```{r}
#Linearity Assumption
plot(interaction_model,1)

#Normality Assumption
plot(interaction_model,2)

#Homoscedascity
plot(interaction_model,3)

#Influencial Outliers
plot(interaction_model,5)

y<-cooks.distance(interaction_model)

summary(y)

```


## Model 1.2 - Interactionand and variable transformation

Building multiple linear regression model with all the variables and relevant interaction term and variables transformation. 
```{r}
options(scipen = 999)
interaction_model_v2 <- lm(a_netm_mon ~ log(a_surface_living) + age + a_sicht + a_balkon + region + a_zip_2 + a_nb_rooms +a_kat_o_2+ region:a_nb_rooms + region:a_sicht + region:a_balkon + region:a_surface_living, data=rent_data)
summary(interaction_model_v2)
```



## Verifying regression assumptions (2)

```{r}
#Linearity Assumption
plot(interaction_model_v2,1)

#Normality Assumption
plot(interaction_model_v2,2)

#Homoscedascity
plot(interaction_model_v2,3)

#Influencial Outliers
plot(interaction_model_v2,5)

y_interaction_model_v2<-cooks.distance(interaction_model_v2)

summary(y_interaction_model_v2)
```

Checking again for correlation between variables, since the introduction of interaction terms seems to have brought multicollinearity. Multicollinarity is present due to interaction terms.
````{r}
library(ggplot2)
library(reshape2)

# Create a matrix of predictors using the formula and data frame
X <- model.matrix(interaction_model_v2, data = rent_data)

# Create a correlation matrix for the variables in the model
corr_matrix <- cor(X)

# Melt the correlation matrix to create a data frame suitable for ggplot2
corr_df <- melt(corr_matrix)

# Create a heatmap of the pairwise correlations
ggplot(corr_df, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Pairwise Correlation Heatmap", x = "", y = "")

````

Appending prediction to original data frame in order to visualise extreme rents data points.
```{r, include=FALSE}
rents_preditions = rent_data %>%
  mutate(predictions = predict(interaction_model, newdata = rent_data))
```


## Model 2 - Stepwise Regression

Stepwise regression model

```{r}
library(MASS)
model_step <- stepAIC(interaction_model_v2, direction = "backward", k = 2, trace = 2)
plot(model_step$anova$AIC, type = "b", xlab = "Number of Variables", ylab = "AIC")


```



The stepwise approach improves the AIC of the model. The same robustness test is performed. 
```{r}
#Linearity Assumption
plot(model_step,1)

#Normality Assumption
plot(model_step,2)

#Homoscedascity
plot(model_step,3)

#Influencial Outliers
plot(model_step,5)

y_model_step<-cooks.distance(model_step)

summary(y_model_step)
````


Looking at the number of observation for each number of rooms. 
```{r}
rent_data %>%
  count(a_nb_rooms)
```

Comparing the regression model (interaction_model_v2) with interaction terms and variable transformation ti the stepwise regression model.
```{r}



#Calculate RSNE
aic_interaction_model_v2 <- AIC(interaction_model_v2)
aic_stepwise <- AIC(model_step)
cat("AIC for model 1.2:", aic_interaction_model_v2, "\n")
cat("AIC for the stepwise regression model:", aic_stepwise, "\n")

#Check length in order to avoid warning (data consitency check)
predicted_interaction_model_v2 <- predict(interaction_model_v2)
predicted_stepwise <- predict(model_step)
actual <- rent_data$a_netm_mon
cat("Length of actual values:", length(actual), "\n")
cat("Length of model 1.2 values:", length(predicted_interaction_model_v2), "\n")
cat("Length of predicted_stepwise values:", length(predicted_stepwise), "\n")

# Calculate RMSE 
rmse_interaction_model_v2 <- suppressWarnings(sqrt(mean((predicted_interaction_model_v2 - actual)^2)))
rmse_stepwise <- suppressWarnings(sqrt(mean((predicted_stepwise - actual)^2)))
cat("RMSE for for model 1.2l:", rmse_interaction_model_v2, "\n")
cat("RMSE for the stepwise regression model:", rmse_stepwise, "\n")


```

## Model 3 - Ridge Regression

Building a ridge regression model

```{r}
#adding variables to formula variable
formula <- a_netm_mon ~ a_surface_living + age + a_sicht + a_balkon + region + a_zip_2 + a_nb_rooms + region:a_sicht + region:a_balkon + region:a_surface_living + region:a_nb_rooms + region:age

#creating a model matrix, it can be used to predict the variable y
X <- model.matrix(formula, data = rent_data)
Y <- rent_data$a_netm_mon

#To verify that values are numeric (optional)
#View(X)

ridge_model<-glmnet(X, Y,alpha = 0)
summary(ridge_model)

#performing cross validation, 10 fold by default
#Y axis is the mean square error, the goal of the regression is to reduce it. To accomplish this different folds are fitted to the data in a cross validation. The graph shows for which value of lambda the mean square value will be minimum. The whole process verifies the robustness of the fitted model.
cv_model<-cv.glmnet(X, Y,alpha = 0)
plot(cv_model)
cv_model$lambda.min

#get the range of lambda values within one standard error of the minimum
lambda_range <- cv_model$lambda.1se

#gheck the range of lambda values
print(cv_model$lambda.1se)
print("sep")
print(cv_model$lambda.min)
#--
#setting lambda in the model
fitted_ridge_model<-glmnet(X, Y,alpha = 0,lambda = cv_model$lambda.min)
coef(fitted_ridge_model)

ypred_ridge <- predict(fitted_ridge_model, newx = X)

RMSE_ridge <- sqrt(mean((Y - ypred_ridge)^2))
print(paste("RMSE for the ridge regression model:", RMSE_ridge))

#calculate R-squared
TSS <- sum((Y - mean(Y))^2) # total sum of squares
RSS <- sum((Y - ypred_ridge)^2) # residual sum of squares

R_squared_ridge <- 1 - (RSS / TSS)

print(paste("R-squared for the ridge regression model:", R_squared_ridge))


```
## Visualization of interaction effects with 0.05 significance level

**Note**: The following plots have been esthitically mdified for the thesis in order to meet standards. The fine design tunning seemed unnessesarily detailed, thus only simple tables are included.

Visualization of the effect of region on the view.
```{r}


# New coefficients
coeff <- c(215.80580268, # German region - View mountain and lake
        -190.97553010, # Italian region - View on mountain and lake

       48.91806510,   # German region - Not mountain and lake
       145.35264984,  # Italian region - Not mountain and lake

       147.28729150,  # German region - View on the lake
       -95.86363811,  # Italian region - View on the lake

       34.24046513,   # German region - View on the mountains
       51.33854512,  # Italian region - View on the mountains

       -65.00933841,  # German region - No view
       0             # Italian region - No view NO DATA
)




Variable <- c("German region - View mountain and lake",
           "Italian region - View on mountain and lake",
           
           "German region - Not mountain and lake",
           "Italian region - Not mountain and lake",
           
           "German region - View on the lake",
           "Italian region - View on the lake",
           
           "German region - View on the mountains",
           "Italian region - View on the mountains",
           
           "German region - No view",
           "Italian region - No view NO DATA"
)
dt<-data.frame(Variable,coeff)

# Plot
library(ggplot2)
dt %>% 
  mutate(Variable = factor(Variable, levels = Variable)) %>%
  ggplot(aes(x = Variable, y = coeff)) +
  geom_bar(stat = "identity", position = "dodge", fill = "black") +
  coord_flip() +
  labs(title = "Comparison of Real Estate Value Based on Views", 
       x = "Location - Scenic View Category",
       y = "Estimated Coefficients") +
  theme_bw() +
  theme(plot.title = element_text(face="bold"), 
        axis.title.x = element_text(face="bold"), 
        axis.title.y = element_text(face="bold"))

```

Visualization of the effect of region on balconies.

```{r}
# New coefficients
coeff <- c(-110.98359471, # German region - General balcony
253.17064821, # Italian region - General balcony

       -133.88474361,  # German region - No balcony
       6.79253953,     # Italian region - No balcony
       
       -67.88322696,   # German region - Roof terasse
       0,              # Italian region - Roof terasse (No data available, kept old value)
       
       -206.42944207,  # German region - Small balcony
       -5.88513478,    # Italian region - Small balcony
       
       42.30112387,    # German region - Terasse balcony
       151.56251002    # Italian region - Terasse balcony
)


Variable <- c("German region - General balcony",
           "Italian region - General balcony",
           
           "German region - No balcony",
           "Italian region - No balcony",
           
           "German region - Roof terasse",
           "Italian region - Roof terasse NO DATA",
           
           "German region - Small balcony",
           "Italian region - Small balcony",
           
           "German region - Terasse balcony",
           "Italian region - Terasse balcony"
)


dt<-data.frame(Variable,coeff)

# Plot
library(ggplot2)
dt %>% 
  mutate(Variable = factor(Variable, levels = Variable)) %>%
  ggplot(aes(x = Variable, y = coeff)) +
  geom_bar(stat = "identity", position = "dodge", fill = "black") +
  coord_flip() +
  labs(title = "Comparison of Real Estate Value Based on Balconies", 
       x = "Type of balcony",
       y = "Estimated Coefficients") +
  theme_bw() +
  theme(plot.title = element_text(face="bold"), 
        axis.title.x = element_text(face="bold"), 
        axis.title.y = element_text(face="bold"))

```

Visualization of the effect of region on the different number of rooms.

```{r}
region_german <- list(
a_nb_rooms1.5 = 51.72145158,
a_nb_rooms2 = 74.50276965,
a_nb_rooms2.5 = 42.60221411,
a_nb_rooms3 = 59.68621761,
a_nb_rooms3.5 = 21.63586165,
a_nb_rooms4 = -28.43466892,
a_nb_rooms4.5 = -61.36824381,
a_nb_rooms5 = -97.47528706,
a_nb_rooms5.5 = -123.97751323,
a_nb_rooms6 = -216.62710495,
a_nb_rooms6.5 = -228.05071391,
a_nb_rooms7 = -338.47081685,
a_nb_rooms7.5 = -224.54492745,
a_nb_rooms8 = -287.21374437,
a_nb_rooms8.5 = 40.82258686,
a_nb_rooms9 = -589.46113411,
a_nb_rooms9.5 = 192.85819822,
a_nb_rooms10 = -101.74969288,
a_nb_rooms10.5 = -278.23149543,
a_nb_rooms14 = 0
)



# Convert the list to a data frame
region_german_df <- data.frame(
  nb_rooms = c(1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.5, 8, 8.5, 9, 9.5, 10, 10.5, 14),
  values = unlist(region_german)
)
# Transforming the data into a data frame
region_german_df <- data.frame(nb_rooms = names(region_german), values = unlist(region_german))

region_german_df %>% 
  mutate(nb_rooms = factor(nb_rooms, levels = nb_rooms)) %>%
  ggplot(aes(x = nb_rooms, y = values)) +
  geom_bar(stat = "identity", position = "dodge", fill = "black") +
  coord_flip() +
  labs(title = "Interaction coefficient of German speaking region", 
       x = "Number of rooms",
       y = "Coefficient") +
  theme_bw() +
  theme(plot.title = element_text(face="bold"), 
        axis.title.x = element_text(face="bold"), 
        axis.title.y = element_text(face="bold"))

```

```{r}


region_italian <- list(
 a_nb_rooms1.5 = -221.78025803,
a_nb_rooms2 = -256.51529117,
a_nb_rooms2.5 = -111.81994498,
a_nb_rooms3 = -35.27375170,
a_nb_rooms3.5 = -20.81747353,
a_nb_rooms4 = -55.88642957,
a_nb_rooms4.5 = -104.22143773,
a_nb_rooms5 = -46.95849964,
a_nb_rooms5.5 = -34.91317283,
a_nb_rooms6 = -410.03982875,
a_nb_rooms6.5 = -316.42893754,
a_nb_rooms7 = -284.54040086,
a_nb_rooms7.5 = -49.95527093,
  a_nb_rooms8 = 0,  # No update provided for a_nb_rooms8
  a_nb_rooms8.5 = 440.80223928,
  a_nb_rooms9 = 0,  # No update provided for a_nb_rooms9
  a_nb_rooms9.5 = 0,  # No update provided for a_nb_rooms9.5
  a_nb_rooms10 = 0,  # No update provided for a_nb_rooms10
  a_nb_rooms10.5 = 0,  # No update provided for a_nb_rooms10.5
  a_nb_rooms14 = 0  # No update provided for a_nb_rooms14
)


# Convert the list to a data frame
region_german_df <- data.frame(
  nb_rooms = c(1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.5, 8, 8.5, 9, 9.5, 10, 10.5, 14),
  values = unlist(region_italian)
)
# Transforming the data into a data frame
region_italian_df <- data.frame(nb_rooms = names(region_italian), values = unlist(region_italian))

region_italian_df %>% 
  mutate(nb_rooms = factor(nb_rooms, levels = nb_rooms)) %>%
  ggplot(aes(x = nb_rooms, y = values)) +
  geom_bar(stat = "identity", position = "dodge", fill = "black") +
  coord_flip() +
  labs(title = "Interaction coefficient of Italian speaking region", 
       x = "Number of rooms",
       y = "Coefficient") +
  theme_bw() +
  theme(plot.title = element_text(face="bold"), 
        axis.title.x = element_text(face="bold"), 
        axis.title.y = element_text(face="bold"))


```




